---
title: "week 5 homework"
output: html_document
---

```{r setup, include=FALSE, purl=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This homework contains some data adapted from Mamerow and colleagues (2016), article to be found here:[https://pubmed.ncbi.nlm.nih.gov/27684105/]. 

The study looks at risk-taking across the age span amongst 902 people aged 18-90, and looks at risk-taking in 2 different ways.

The self.csv file contains a self-reported risk taking measure, as well as background variables. Variables comprise:

* partid: participant id
* sex: (0=male, 1=female)
* age: age in years
* riskSR: self reported risk attitude  (0-10)

The gambles.csv file contains experimental data for the same participants when they performed a gambling simulation. On each trial the participant makes a forced choice decision between 2 investments, each with a specific pay-off and specific chance of success. Each time there is a risky and a safe choice, differing in profile. Any money earned is kept by the participant to incentivise them to choose realistically.

Each person did 8 trials, split into 4 'normal' and 4 'higher risk' trials, where taking the risky option would lead to a loss in the long run. The variables are:

* partid: participant id
* scenario: scenario of the gamble (there were 8 scenarios, each person saw each scenario once)
* trialtype: 0 = normal trial, 1 = higher risk trial
* RiskyChoice: 0 = safe choice chosen, 1 = risky choice chosen

First load the data and any libraries you will need.
```{r purl=FALSE}
self <- read.csv('self.csv') 

gambles <- read.csv('gambles3.csv') 
library(tidyverse)
library(lmerTest)

```


Q1. We want to predict if self reported risk attitude is associated with age. What type of model do we need for this? What is the DV? Are there any random effects?
```{r q01}
#linear regression
#DV: self-reported risk attitude (numeric)
#random effects: participant id, sex
```

Q2. First we will centre and scale the age variable in the `self` dataframe and call it `age_cs` Note: this is what they did in the original study (unlike previous times when we have merely centrered it).
```{r q02}
self<- self %>% mutate(age_cs = scale(age))
```


Q3. Plot the data showing how age predicts self reported risk taking, with a different regression line for each sex

HINT: what data type does sex need to be for this to work?
```{r q03}
ggplot(self, aes(x = age_cs, y = riskSR, group = sex %>% as.factor()))+
  geom_point()+
  geom_smooth(method = "lm",aes(color=sex %>% as.factor()))
```

Q4. Find out for yourself: Try to improve the layout of the plot so that the legend informs us which line is which. HINT the layer starts with `scale_` and contains the aes mapping and the data type.

```{r q04}
ggplot(self, aes(x = age_cs, y = riskSR))+
  geom_point()+
  geom_smooth(method = "lm",aes(color=sex %>% as.factor()))+
  scale_color_discrete(name = "Sex", labels = c("Male", "Female"))
```

Q5. What does the graph suggest about risk attitudes
```{r q05}
#risk attitudes decline with age, males have a higher intercept and slope than females. But the correlation is not strong.
```


Q6. This code is to run the model with age_cs as a predictor and call it mod1. In the paper they included linear, quadratic and cubic effects of age.
Can you fix the code so that it runs?
```{r q06}
mod1 <- lm(riskSR ~ age_cs + I(age_cs^2) + I(age_cs^3), data=self)
summary(mod1)
```

Q7. Does adding an effect of sex improve the model - call this `mod2`? 
```{r q07}
mod2 <- lm(riskSR ~ age_cs + I(age_cs^2) + I(age_cs^3) + sex, data=self)
summary(mod2)
anova(mod1,mod2)
#adding an effect of sex improves the model, p<.001
```


Q8. Now lets combine the self and gambles data and call this `combi`
```{r q08}
combi<- inner_join(self,gambles, by = "partid")
```

Q9. Overall, on what proportion of trials did people made the riskier choice under the 'normal' and 'high risk' conditions
```{r q09}
combi%>%group_by(trialtype, RiskyChoice)%>%count()%>% mutate(percent = n/14401)
#normal 23.96%
#high risk 19.37%
```

Q10. How do you interpret this in words?
```{r q10}
#when trials are more risky (trialtype=1) the tendency to choose the safer option increases
```

Q11. We want to model RiskyChoice as the dependent variable, with trialtype and age as predictors. What random effects should we consider?
```{r q11}
mod3<- glmer(RiskyChoice ~ trialtype + age + (1|partid) + (1|scenario), data = combi)
summary(mod3)
#we need to control for partid and scenario
```


Q12. Build a model with a fixed effect of trial type - include random intercepts and slopes for participant and random intercepts for scenarios and call it `mod4`. 

```{r q12}
mod4<- glmer(RiskyChoice ~ trialtype +  (1+trialtype|partid) + (1|scenario), data = combi)
summary(mod4)
```

Q13. The final question involves using Rstudio to push a file to GitHub.

Navigate to this repo: https://github.com/PALS0043/Week5Homework
Hit the green code button and copy the https code to clone it.
Open a new project in RStudio linked to this repo.
Copy this homework file to the project folder on your computer.
Use the git pane to commit and push the homework file to GitHub.
NB: please ensure you have named the file with your candidate number.

